% Market Event Radar PRD v1.1 (LaTeX master copy)
% Compile: pdflatex market_event_radar_prd_v1.1.tex (run twice for TOC)
\documentclass[11pt]{article}

\usepackage[letterpaper, margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{helvet}
\renewcommand{\familydefault}{\sfdefault}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{array}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{parskip}
\usepackage[colorlinks=true, linkcolor=blue!60!black]{hyperref}

\definecolor{prdblue}{HTML}{1F3864}
\definecolor{prdaccent}{HTML}{2E5395}
\definecolor{calloutbg}{HTML}{EEF3FA}
\definecolor{warnbg}{HTML}{FDEBD0}

\usepackage{sectsty}
\allsectionsfont{\color{prdblue}}

\usepackage[most]{tcolorbox}
\newtcolorbox{callout}[1]{colback=calloutbg, colframe=prdaccent, boxrule=0.5pt,
  arc=1mm, left=2mm, right=2mm, top=1.5mm, bottom=1.5mm, title=\textbf{#1},
  coltitle=black, colbacktitle=calloutbg, attach title to upper={\par}}
\newtcolorbox{warncallout}[1]{colback=warnbg, colframe=orange!70!black, boxrule=0.5pt,
  arc=1mm, left=2mm, right=2mm, top=1.5mm, bottom=1.5mm, title=\textbf{#1},
  coltitle=black, colbacktitle=warnbg, attach title to upper={\par}}

\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\setlist[itemize]{topsep=2pt, itemsep=2pt, parsep=0pt, leftmargin=1.5em}
\setlist[enumerate]{topsep=2pt, itemsep=2pt, parsep=0pt, leftmargin=1.8em}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small Market Event Radar \,|\, PRD v1.1}
\fancyhead[R]{\small\textbf{CONFIDENTIAL}}
\fancyfoot[L]{\small Market Event Radar, Weekly Catalyst Digest}
\fancyfoot[R]{\small Page \thepage\ of \pageref{LastPage}}
\renewcommand{\headrulewidth}{0.4pt}

\title{\vspace{-1.2em}{\Huge\bfseries\color{prdblue} MARKET EVENT RADAR}\\[0.3em]
{\large\itshape Weekly Catalyst Digest, Product Requirements Document}\\[0.5em]
{\normalsize Version 1.1 \,|\, June 2026 \,|\, US Equity Markets (NYSE, NASDAQ)}}
\author{}
\date{}

\begin{document}

\maketitle
\vspace{-3em}

\begin{center}
\begin{tabular}{|l|l|l|l|}
\hline
\textbf{Status} & \textbf{Author} & \textbf{Market} & \textbf{Stack} \\
\hline
Draft v1.1 & Shreyas & NYSE / NASDAQ & Python, Free Only \\
\hline
\end{tabular}

\vspace{0.8em}
\textbf{CONFIDENTIAL, INTERNAL USE ONLY}
\end{center}

\tableofcontents
\newpage

%=====================================================================
\section{Overview}

Market Event Radar is a free, automated daily digest tool that surfaces high-impact events happening in the next 7 days across US-listed equities (NYSE and NASDAQ). It helps an individual investor or trader start each morning knowing exactly which stocks are about to face potential price-moving catalysts, without subscribing to any paid data service.

The tool focuses exclusively on stocks with active options markets, ensuring every name in the digest has sufficient liquidity and institutional attention to be worth monitoring.

\begin{callout}{Core Product Principle: Events, Not Predictions}
This tool reports factual, verifiable event information only. It does not predict price direction, generate buy/sell signals, or label stocks bullish or bearish. All options-activity data is presented as raw, descriptive market facts (volume, open interest, put/call ratio) with no directional interpretation.
\end{callout}

%=====================================================================
\section{Problem Statement}

Market-moving events (earnings reports, IPO listings, M\&A announcements, legal rulings) are publicly schedulable, but they are scattered across dozens of sources: SEC EDGAR, financial news wires, earnings calendars, court dockets, and analyst feeds. A retail trader has no single, free, consolidated view of what is coming this week.

Existing solutions are either expensive (Bloomberg, FactSet), noisy (generic news feeds with no event structure), or limited in scope (earnings-only calendars). Additionally, no single free data source is complete on its own: free providers routinely skip tickers, lag on updates, or cover only large, well-known names. This project fills the gap with a zero-cost, structured, daily digest built on redundant, cross-checked free sources.

%=====================================================================
\section{Goals \& Non-Goals}

\subsection{Goals}
\begin{itemize}
  \item \textbf{Consolidated weekly event view:} Show all material upcoming events for the next 7 days in a single daily digest.
  \item \textbf{Free-only data sources:} Zero API subscription cost. All data sourced from free tiers or open APIs.
  \item \textbf{No single-source dependency:} Every critical data type (earnings dates, options data, market cap) is fetched from at least two independent free sources and merged, so a gap or outage in one source does not silently drop events.
  \item \textbf{Options-volume gated:} Only show stocks with active, liquid options markets, no thinly traded names.
  \item \textbf{Multi-category coverage:} Earnings reports, new IPO listings, SEC 8-K filings (M\&A, legal), and news catalysts.
  \item \textbf{Market-cap classification:} Tag each stock as Large Cap, Mid Cap, or Small Cap for easy filtering.
  \item \textbf{Actionable output:} A clean daily HTML digest the user can open each morning, ranked by event significance and options activity.
\end{itemize}

\subsection{Non-Goals}
\begin{itemize}
  \item No price prediction of any kind. The tool never states or implies whether a stock will go up or down.
  \item No directional labels. Options positioning data (put/call ratio, volume spikes) is shown as raw numbers only, never translated into bullish/bearish sentiment tags.
  \item No trade recommendations or buy/sell signals.
  \item No real-time alerts or push notifications (Phase 1 scope is daily batch).
  \item No coverage of non-US exchanges (India, Europe, etc.) in v1.
  \item No paid APIs, subscriptions, or data licenses of any kind.
  \item No proprietary order-flow or dark-pool data.
\end{itemize}

%=====================================================================
\section{User Persona}

\begin{tabularx}{\textwidth}{|L{3.2cm}|X|}
\hline
\textbf{Name} & Shreyas, Retail Investor / Swing Trader \\
\hline
\textbf{Goal} & Know what events are coming this week before the market opens \\
\hline
\textbf{Workflow} & Opens the digest each morning, scans for names on his watchlist or by cap tier, decides whether to enter or avoid a position \\
\hline
\textbf{Pain Point} & Currently checks 4-5 separate sites manually (earnings whispers, SEC EDGAR, financial Twitter, a news app). Misses events. No options filter exists anywhere for free. \\
\hline
\textbf{Technical Level} & Comfortable running a Python script. Does not want to manage paid APIs or complex infra. \\
\hline
\end{tabularx}

%=====================================================================
\section{Ticker Universe Definition \hfill {\small [NEW in v1.1]}}

The candidate ticker universe for each daily run is constructed dynamically as the union of:

\begin{itemize}
  \item \textbf{Earnings calendar tickers:} All NYSE/NASDAQ tickers with earnings scheduled in the next 7 days (merged across sources, see Section~\ref{sec:sources}).
  \item \textbf{IPO calendar entries:} Companies pricing or listing within the next 7 days (Phase 2).
  \item \textbf{8-K filers:} All companies that filed a monitored 8-K item with the SEC in the last 24 hours.
  \item \textbf{User watchlist:} Tickers listed in a local config file (\texttt{tickers.txt} / \texttt{config.yaml}). Watchlist tickers are always evaluated, every run, regardless of whether a calendar source surfaced them. This is the user's safety net against source coverage gaps.
\end{itemize}

The union is deduplicated, restricted to NYSE/NASDAQ common stock (no OTC, no ADRs, no ETFs), then passed through the options volume gate (Section~\ref{sec:optionsfilter}). News catalyst searches (Section~\ref{sec:news}) run only against tickers already in this universe, keeping news API usage bounded.

%=====================================================================
\section{Event Categories}

The digest covers four primary categories of market-moving events. Each maps to specific data sources and has a defined priority weight in the output ranking.

\subsection{Earnings Reports}
Scheduled quarterly earnings releases where a company reports revenue, EPS, and guidance. These are the single highest-impact, most predictable event type for options traders.
\begin{itemize}
  \item Coverage window: Earnings due within the next 7 calendar days
  \item Key fields: Ticker, Company Name, Report Date, Report Time (BMO / AMC), EPS Estimate, Revenue Estimate, Confirmation Status
  \item \textbf{Confirmation status [NEW in v1.1]:} Each earnings entry is tagged CONFIRMED (company has announced the date) or ESTIMATED (provider projection). Estimated dates are a major noise source; the digest visually distinguishes them so the user knows which dates can move.
  \item \textbf{Multi-source merge [NEW in v1.1]:} Earnings calendar is the union of Finnhub and yfinance results, cross-checked against Alpha Vantage when quota allows. See Section~\ref{sec:chains} for reconciliation rules.
  \item Priority weight: HIGH
\end{itemize}

\subsection{IPO \& New Listings}
Companies going public for the first time (IPO) or uplisting from OTC to NYSE/NASDAQ. These create sudden liquidity events and often have extreme first-day price swings.
\begin{itemize}
  \item Coverage window: IPOs priced or listing within the next 7 days
  \item Key fields: Company Name, Expected Listing Date, Exchange, Estimated Price Range, Sector
  \item Sources: Finnhub IPO calendar, cross-checked against SEC EDGAR S-1/424B filings
  \item Note: IPOs may not yet have an options chain, flagged separately and exempted from the options gate
  \item Priority weight: HIGH
\end{itemize}

\subsection{SEC 8-K Filings, Material Events}
Companies must file an 8-K with the SEC within 4 business days of any material event. This includes M\&A agreements, major contracts, leadership changes, asset sales, and legal settlements. These are legally required disclosures and are highly reliable signals.
\begin{itemize}
  \item Coverage window: 8-K filings from the last 24 hours that reference forward-dated events within 7 days
  \item Key item codes to monitor: 1.01 (Material Agreements), 1.02 (Terminations), 1.03 (Bankruptcy), 2.01 (Asset Acquisitions/Disposals), 5.02 (Executive Changes), 8.01 (Other Events)
  \item Source: SEC EDGAR is authoritative and primary; Finnhub filing summaries serve as a secondary feed to catch EDGAR fetch failures
  \item Priority weight: MEDIUM-HIGH (depends on item type)
\end{itemize}

\subsection{News Catalysts}\label{sec:news}
Breaking or anticipated news that could influence stock price but is not yet reflected in a scheduled event. Examples: FDA decision dates, regulatory rulings, contract award announcements, court hearing outcomes.
\begin{itemize}
  \item Coverage window: News articles dated within the last 48 hours referencing events in the next 7 days
  \item Sources: Finnhub company news (primary), NewsAPI keyword search (secondary), Alpha Vantage news feed (tertiary, quota permitting)
  \item Filtered by: Keyword scoring (merger, acquisition, ruling, settlement, approval, award, verdict, etc.)
  \item Scope: News lookups run only against tickers already in the universe (Section 5)
  \item Priority weight: MEDIUM (signal quality varies)
\end{itemize}

%=====================================================================
\section{Options Volume Filter}\label{sec:optionsfilter}

\begin{callout}{Core Design Principle}
Every stock in the digest must pass the options volume gate. A stock with no active options market is not actionable for most traders and is excluded entirely. (Exception: IPOs without a chain yet, flagged separately.)
\end{callout}

The filter operates in two layers:

\subsection{Baseline Liquidity Gate (Pass / Fail)}
A stock must meet ALL of the following criteria to appear in the digest:

\begin{tabularx}{\textwidth}{|L{4.6cm}|L{3.8cm}|X|}
\hline
\textbf{Criterion} & \textbf{Threshold} & \textbf{Rationale} \\
\hline
Has options chain & Yes (any expiry) & Confirms options market exists \\
\hline
Total open interest & $>$ 1{,}000 contracts & Minimum institutional interest \\
\hline
Avg daily options volume (20d) & $>$ 500 contracts/day (per-tier thresholds under evaluation, see Open Questions) & Ensures tradeable liquidity \\
\hline
Avg daily stock volume (20d) & $>$ 500{,}000 shares/day & Basic equity liquidity check \\
\hline
\end{tabularx}

\textbf{Source redundancy [NEW in v1.1]:} Options chain data is fetched from yfinance (primary) with Polygon.io free tier as fallback. If both fail for a ticker, the ticker is marked DATA UNAVAILABLE rather than silently dropped, and appears in a diagnostics footer of the digest.

\subsection{Elevated Options Activity (Descriptive Only) \hfill {\small [REVISED in v1.1]}}
After passing the gate, each stock's current options volume is compared against its own 20-day average. Elevated activity means more market participants are positioned in the name ahead of a known event; the digest reports this purely as a factual liquidity observation and uses it for ranking order only. No directional meaning is assigned.

\begin{tabularx}{\textwidth}{|L{3.4cm}|X|L{3.6cm}|}
\hline
\textbf{Signal} & \textbf{Definition} & \textbf{Digest Rank Boost} \\
\hline
Elevated, Strong & Today's options vol $>$ 3x 20-day avg & +2 rank positions \\
\hline
Elevated, Moderate & Today's options vol $>$ 1.5x 20-day avg & +1 rank position \\
\hline
Put/Call Ratio & Raw ratio displayed as a number (e.g., P/C: 0.62). No bullish/bearish label is applied. & None (informational only) \\
\hline
\end{tabularx}

\textit{Removed in v1.1: the Call/Put Skew and Put Skew rows that labeled stocks with a bullish or bearish lean. Directional interpretation conflicts with the events-only product principle (Section 1). The underlying put/call ratio remains visible as a raw figure for users who want to interpret it themselves.}

%=====================================================================
\section{Market Cap Classification}

Every stock in the digest is tagged with its cap tier. This helps the user quickly filter to their preferred risk profile (large caps are more predictable; small caps can have wilder swings).

\begin{tabularx}{\textwidth}{|L{2.4cm}|L{3.4cm}|X|L{2.4cm}|}
\hline
\textbf{Cap Tier} & \textbf{Market Cap Range} & \textbf{Typical Options Liquidity} & \textbf{Digest Tag} \\
\hline
Large Cap & $>$ \$10 Billion & High, weekly/daily chains & LARGE \\
\hline
Mid Cap & \$2B - \$10B & Moderate, monthly chains & MID \\
\hline
Small Cap & \$300M - \$2B & Thin, may fail OI gate & SMALL \\
\hline
Micro Cap & $<$ \$300M & Excluded by options gate & EXCLUDED \\
\hline
\end{tabularx}

\textit{Market cap is sourced from yfinance (primary) with Finnhub company profile as fallback (v1.1). Classification is applied after the options gate; micro caps almost always fail the gate and are excluded automatically.}

%=====================================================================
\section{Data Sources \& Redundancy Strategy \hfill {\small [EXPANDED in v1.1]}}\label{sec:sources}

\begin{callout}{Constraint}
No paid APIs, subscriptions, or data licenses are permitted in this project. All sources below are free-tier or open-access.
\end{callout}

\begin{warncallout}{Redundancy Principle [NEW in v1.1]}
No critical data type may depend on a single source. Free providers have incomplete coverage: yfinance can skip tickers or return data only for widely held names, Finnhub free tier truncates some calendars, and any source can be temporarily down. Each data type therefore has a defined primary source plus at least one fallback, and calendar-type data is merged as a union across sources, not taken from whichever source answers first.
\end{warncallout}

\subsection{Source Inventory}

\begin{tabularx}{\textwidth}{|L{2.1cm}|X|L{1.9cm}|L{2.9cm}|L{1.4cm}|}
\hline
\textbf{Source} & \textbf{Data Provided} & \textbf{Auth} & \textbf{Rate Limit} & \textbf{Cost} \\
\hline
yfinance & Earnings dates, options chain (vol, OI), market cap, 20d avg volume, price history & None (Python lib) & Unofficial, generous & Free \\
\hline
Finnhub & Earnings calendar, IPO calendar, company news, filing summaries, company profile & Free API key & 60 calls/min & Free tier \\
\hline
SEC EDGAR & Full 8-K filings (M\&A, legal, executive changes), S-1/IPO filings, full-text search & None (open REST API) & 10 req/sec & Free \\
\hline
Alpha Vantage & News \& sentiment feed, earnings history/calendar & Free API key & 5 calls/min, 500/day & Free tier \\
\hline
Polygon.io & Options snapshot (15-min delayed), OI, volume by strike & Free API key & 5 calls/min on free tier (verify before build) & Free tier \\
\hline
NewsAPI & Broad financial news keyword search & Free API key & 100 req/day (24h delay) & Free tier \\
\hline
\end{tabularx}

\textit{Correction in v1.1: Polygon.io free tier was previously listed as unlimited; it is rate-limited (approximately 5 calls/min). It is therefore demoted from batch options snapshot provider to per-ticker fallback only.}

\subsection{Primary / Fallback Chains per Data Type}\label{sec:chains}

\begin{tabularx}{\textwidth}{|L{2.9cm}|L{1.9cm}|L{2.6cm}|X|}
\hline
\textbf{Data Type} & \textbf{Primary} & \textbf{Secondary} & \textbf{Merge / Reconciliation Rule} \\
\hline
Earnings calendar & Finnhub & yfinance (+ Alpha Vantage spot-checks) & UNION of all sources. If dates disagree, show earliest date, tag ESTIMATED, list both dates in the event card \\
\hline
IPO calendar & Finnhub & SEC EDGAR S-1 / 424B scan & Union; EDGAR filing confirms Finnhub entries \\
\hline
8-K material events & SEC EDGAR & Finnhub filing feed & EDGAR is authoritative; Finnhub fills gaps only on EDGAR fetch failure \\
\hline
Options chain / OI / volume & yfinance & Polygon.io (delayed) & Primary wins; fallback used per-ticker on failure or empty chain. Both fail: tag DATA UNAVAILABLE, never silently drop \\
\hline
Market cap & yfinance & Finnhub company profile & Primary wins; fallback on missing/zero value \\
\hline
Company news & Finnhub & NewsAPI, Alpha Vantage & Union with URL-based dedup; quota-capped \\
\hline
\end{tabularx}

\begin{itemize}
  \item \textbf{Source attribution:} Every event object carries a \texttt{sources} field (e.g., \texttt{[finnhub, yfinance]}). The digest shows a small source tag per event so the user can judge reliability, and single-source events are visually marked since they lack cross-confirmation.
  \item \textbf{Coverage diff logging:} On every run, the pipeline logs tickers found by one earnings source but not another. A persistently one-sided diff signals a source degrading, surfaced in the digest diagnostics footer.
\end{itemize}

%=====================================================================
\section{Technical Architecture}

\subsection{Pipeline Overview}
The system runs as a single Python script executed once daily (e.g., 6:00 AM ET). It fetches from redundant sources, merges, filters, ranks, and writes an HTML digest file to disk.

\begin{tabularx}{\textwidth}{|L{0.8cm}|L{3.1cm}|X|L{3.0cm}|}
\hline
\textbf{Step} & \textbf{Module} & \textbf{Action} & \textbf{Output} \\
\hline
1 & \texttt{event\_fetcher.py} & Pull earnings calendar (Finnhub + yfinance), IPO calendar (Finnhub + EDGAR), 8-K filings (EDGAR, Finnhub fallback) for next 7 days; include watchlist tickers & Raw event lists per source \\
\hline
2 & \texttt{source\_merger.py} [NEW] & Union-merge events across sources, dedup by (ticker, event type, date window), apply reconciliation rules (Sec.~\ref{sec:chains}), attach \texttt{sources} field & Unified event list \\
\hline
3 & \texttt{options\_filter.py} & Per ticker: fetch options chain (yfinance, Polygon fallback), check OI $>$ 1000, avg vol gate. Compute elevated-activity ratio and raw put/call ratio & Filtered tickers + activity metrics \\
\hline
4 & \texttt{enricher.py} & Fetch market cap (yfinance, Finnhub fallback), classify cap tier, fetch news (Finnhub/NewsAPI), parse 8-K item codes, set earnings confirmation status & Enriched event objects \\
\hline
5 & \texttt{dedup\_store.py} [NEW] & Compare against \texttt{seen\_events.json}; tag repeats as PREVIOUSLY REPORTED or suppress per config; persist today's event hashes & Deduplicated event list \\
\hline
6 & \texttt{ranker.py} & Score events: event type weight + activity boost + days until event (closer = higher) + cap weight & Ranked event list \\
\hline
7 & \texttt{digest\_writer.py} & Generate HTML digest with today's date. Group by event type. Tag cap tier, confirmation status, sources. Append diagnostics footer & \texttt{market\_radar\_YYYY-MM-DD.html} \\
\hline
\end{tabularx}

\subsection{Technology Stack}

\begin{tabularx}{\textwidth}{|L{3.6cm}|X|}
\hline
\textbf{Component} & \textbf{Choice} \\
\hline
Language & Python 3.10+ \\
\hline
Core data library & yfinance (\texttt{pip install yfinance}), no API key, completely free \\
\hline
HTTP requests & \texttt{requests} library for Finnhub, EDGAR, Alpha Vantage, Polygon, NewsAPI \\
\hline
Scheduling & cron (macOS/Linux) or Task Scheduler (Windows), Phase 2 \\
\hline
Output format & Self-contained HTML file with inline CSS, opens in any browser \\
\hline
Storage & Local filesystem: one HTML file per day, plus \texttt{seen\_events.json} and \texttt{cache/} directory. No database \\
\hline
\end{tabularx}

\subsection{Rate Limit Management}
\begin{itemize}
  \item \textbf{Finnhub:} 1-second sleep between requests; batch ticker lookups where supported.
  \item \textbf{SEC EDGAR:} Respect the 10 req/sec rule. Send a valid User-Agent header (required by EDGAR policy).
  \item \textbf{Alpha Vantage:} Enrichment fallback only; skip when daily quota is near.
  \item \textbf{Polygon.io:} Per-ticker fallback only, throttled to free-tier limit (about 5 calls/min).
  \item \textbf{yfinance:} Unofficial, no hard limit, but add 0.5s sleep between tickers to avoid IP bans. Caching (Section 11) keeps total call volume low.
\end{itemize}

%=====================================================================
\section{Resilience, Caching \& Deduplication \hfill {\small [NEW in v1.1]}}

\subsection{Graceful Degradation}
A source failure must never kill the whole run or silently shrink the digest.
\begin{itemize}
  \item If a primary source fails, the fallback in Section~\ref{sec:chains} is used automatically.
  \item If all sources for a section fail, the digest still generates, with a visible banner in that section (e.g., \textit{Earnings section unavailable today: all sources failed}) instead of an empty or missing section.
  \item Per-ticker failures are tagged DATA UNAVAILABLE and listed in a diagnostics footer with the failing source and error type.
  \item The script exits 0 with a partial digest; it exits non-zero only if no section could be produced at all.
\end{itemize}

\subsection{Local Caching}
\begin{itemize}
  \item Market cap: cached locally, refreshed weekly (changes slowly).
  \item 20-day average stock/options volume: cached daily per ticker.
  \item Options chain snapshots: cached for the current run only (no staleness across days).
  \item Purpose: cuts yfinance call volume sharply, reduces IP-ban risk, and protects the sub-3-minute runtime target as the universe grows.
\end{itemize}

\subsection{Event Deduplication Across Days}
\begin{itemize}
  \item Each event gets a stable hash: (ticker, event type, event date, source document ID where available).
  \item Hashes are persisted in \texttt{seen\_events.json} (flat file, no database, consistent with the storage constraint).
  \item Repeat events (e.g., the same 8-K appearing in tomorrow's 24h window overlap) are tagged PREVIOUSLY REPORTED, or fully suppressed via a config flag.
  \item \texttt{seen\_events.json} entries expire after 14 days to keep the file small.
\end{itemize}

%=====================================================================
\section{Digest Output Format}

\subsection{Structure}
\begin{itemize}
  \item Header: Date, market summary (SPY/QQQ), total events found, source health summary
  \item Section A: Earnings This Week (sorted by date, then activity score; CONFIRMED vs ESTIMATED visually distinct)
  \item Section B: Upcoming IPOs \& New Listings
  \item Section C: SEC 8-K Material Events (M\&A, Legal, Executive)
  \item Section D: News Catalysts (keyword-matched, scored)
  \item Footer: Diagnostics (per-source status, tickers with DATA UNAVAILABLE, coverage diffs)
\end{itemize}

\subsection{Per-Event Card Fields}

\begin{tabularx}{\textwidth}{|L{3.6cm}|X|}
\hline
\textbf{Field} & \textbf{Description} \\
\hline
Ticker + Company & e.g., AAPL, Apple Inc. \\
\hline
Cap Tag & LARGE / MID / SMALL \\
\hline
Event Type & EARNINGS / IPO / 8-K: M\&A / 8-K: LEGAL / NEWS \\
\hline
Event Date & e.g., Tuesday Jun 17, BMO \\
\hline
Confirmation & CONFIRMED / ESTIMATED \quad [NEW in v1.1, earnings only] \\
\hline
Days Away & e.g., 2 days \\
\hline
Options Activity & Avg vol, OI, elevated-activity flag if today's vol $>$ 1.5x 20d avg \\
\hline
Put/Call Ratio & Raw number only (e.g., P/C: 0.62). No directional label. \quad [REVISED in v1.1] \\
\hline
Sources & Which providers reported the event (e.g., finnhub + yfinance). Single-source events marked. \quad [NEW in v1.1] \\
\hline
Brief Summary & 1-2 sentence plain-language, factual description of the event. No predictive language. \\
\hline
\end{tabularx}

%=====================================================================
\section{Ranking Logic}

Events within each section are ranked by a composite score. Ranking affects display order only; it is a relevance ordering, not a trade signal.

\begin{callout}{Rank Score Formula}
score = (event\_type\_weight $\times$ 10) + (activity\_multiplier $\times$ 5) + (10 $-$ days\_away) + (cap\_weight $\times$ 2)
\end{callout}

\begin{tabularx}{\textwidth}{|X|L{1.6cm}|L{6.2cm}|}
\hline
\textbf{Variable} & \textbf{Value} & \textbf{Notes} \\
\hline
event\_type\_weight (Earnings) & 5 & Highest, most predictable impact \\
\hline
event\_type\_weight (IPO) & 4 & High, first-day volatility \\
\hline
event\_type\_weight (8-K M\&A / Legal) & 3 & Medium-high, material disclosure \\
\hline
event\_type\_weight (News Catalyst) & 2 & Medium, quality varies \\
\hline
activity\_multiplier (strong) & 2.0 & Options vol $>$ 3x 20d avg \\
\hline
activity\_multiplier (moderate) & 1.5 & Options vol $>$ 1.5x 20d avg \\
\hline
activity\_multiplier (normal) & 1.0 & No elevated activity \\
\hline
days\_away & 1-7 & Closer events rank higher (10 $-$ days) \\
\hline
cap\_weight (Large / Mid / Small) & 3 / 2 / 1 & Larger cap = more options liquidity \\
\hline
\end{tabularx}

%=====================================================================
\section{Implementation Plan \hfill {\small [NEW in v1.1]}}\label{sec:impl}

Implementation is divided into four phases. Each phase has concrete steps, deliverables, and an exit criterion that must be met before moving on. Steps within a phase are ordered so that every step produces something runnable and testable on its own.

\subsection{Phase 0: Project Setup \hfill {\small (target: half a day)}}

\begin{enumerate}
  \item \textbf{Repository \& environment:} Create the Git repo with the structure below. Set up Python 3.10+ virtualenv, \texttt{requirements.txt} (yfinance, requests, pyyaml, jinja2 for HTML templating, pytest).
\begin{verbatim}
market-event-radar/
  config.yaml          # API keys (env-var refs), watchlist path,
                       # thresholds, top-N, dedup mode
  tickers.txt          # user watchlist, one ticker per line
  src/
    universe.py        # Phase 1, Step 2
    event_fetcher.py   # Phase 1, Step 3
    source_merger.py   # Phase 1, Step 4
    options_filter.py  # Phase 1, Step 5
    enricher.py        # Phase 1, Step 6
    dedup_store.py     # Phase 1, Step 7
    ranker.py          # Phase 1, Step 8
    digest_writer.py   # Phase 1, Step 9
    cache.py           # Phase 1, Step 5
    sources/           # one thin client per provider
      finnhub_client.py
      yfinance_client.py
      edgar_client.py
      polygon_client.py
  cache/               # gitignored
  output/              # gitignored, daily HTML files
  tests/
  run_daily.py         # orchestrator entry point
\end{verbatim}
  \item \textbf{API keys:} Register free Finnhub and Polygon.io keys. Store as environment variables; \texttt{config.yaml} references variable names only, never raw keys (repo stays safe to publish).
  \item \textbf{Source client skeletons:} Each provider client exposes typed functions (e.g., \texttt{get\_earnings\_calendar(start, end)}) with built-in rate-limit sleeps, retries (3 attempts, exponential backoff), timeout, and a uniform \texttt{SourceResult\{ok, data, error\}} return shape. This uniform shape is what makes Phase 1 fallback logic trivial.
  \item \textbf{EDGAR compliance:} Hardcode the required User-Agent header (name + contact email) in \texttt{edgar\_client.py} from day one.
\end{enumerate}
\textbf{Exit criterion:} \texttt{python -c "import src"} clean; each client can make one successful authenticated call (smoke test script).

\subsection{Phase 1: MVP, Earnings + 8-K Digest \hfill {\small (target: 1.5-2 weeks)}}

\begin{enumerate}
  \item \textbf{Step 1, Config loader:} Parse \texttt{config.yaml} and \texttt{tickers.txt}; validate thresholds and paths; fail fast with clear messages on bad config.
  \item \textbf{Step 2, Universe builder (\texttt{universe.py}):} Implement the Section 5 union (earnings tickers + 8-K filers + watchlist), exchange filter (NYSE/NASDAQ common stock only), dedup. Unit-test with fixture lists.
  \item \textbf{Step 3, Event fetchers (\texttt{event\_fetcher.py}):}
    \begin{itemize}
      \item Earnings via Finnhub (primary) and yfinance (secondary), each returning the common event schema.
      \item 8-K via EDGAR full-text/submissions API filtered to monitored item codes; Finnhub filing feed as fallback.
      \item Each fetcher is independently runnable from the CLI for debugging (e.g., \texttt{python -m src.event\_fetcher earnings}).
    \end{itemize}
  \item \textbf{Step 4, Source merger (\texttt{source\_merger.py}):} Union-merge with dedup key (ticker, event type, date window $\pm$2 days). Apply reconciliation rules from Section~\ref{sec:chains}; attach \texttt{sources} list and CONFIRMED/ESTIMATED status; emit the coverage diff log. Unit-test the disagreement cases explicitly (same ticker, dates 1 day apart vs 3 days apart).
  \item \textbf{Step 5, Options gate + caching (\texttt{options\_filter.py}, \texttt{cache.py}):} Per-ticker chain fetch with yfinance, Polygon fallback, DATA UNAVAILABLE tagging. Compute OI total, 20d averages, elevated-activity ratio, raw put/call ratio. Implement the file-based cache (market cap weekly, averages daily) before scaling to the full universe, in this order, or the first full run risks an IP ban.
  \item \textbf{Step 6, Enricher (\texttt{enricher.py}):} Market cap (yfinance, Finnhub fallback), cap tier classification, 8-K item code parsing into human-readable event subtypes.
  \item \textbf{Step 7, Dedup store (\texttt{dedup\_store.py}):} Stable event hashing, \texttt{seen\_events.json} persistence, 14-day expiry, PREVIOUSLY REPORTED tagging with config-flag suppression.
  \item \textbf{Step 8, Ranker (\texttt{ranker.py}):} Implement the Section 13 formula as a pure function. Unit-test with hand-computed expected scores.
  \item \textbf{Step 9, Digest writer (\texttt{digest\_writer.py}):} Jinja2 HTML template, inline CSS, sections A and C, per-event cards with all v1.1 fields, diagnostics footer, section-failure banners.
  \item \textbf{Step 10, Orchestrator + resilience (\texttt{run\_daily.py}):} Wire steps 1-9 with per-stage try/except implementing graceful degradation (Section 11.1) and structured logging to \texttt{output/run\_YYYY-MM-DD.log}.
  \item \textbf{Step 11, Validation week:} Run manually every morning for 5 trading days. Measure against Section 16 metrics: earnings capture vs the union, runtime, quota usage, noise rate (subjective tally), zero illiquid leaks. Tune thresholds (Open Question 3) from real data.
\end{enumerate}
\textbf{Exit criterion:} 5 consecutive clean daily runs; $>$95\% S\&P 500 earnings capture; runtime $<$ 3 min; one simulated source outage (disable Finnhub key) still produces a partial digest with banner.

\subsection{Phase 2: Coverage Expansion + Automation \hfill {\small (target: 1-1.5 weeks)}}

\begin{enumerate}
  \item \textbf{IPO calendar:} Finnhub IPO endpoint + EDGAR S-1/424B cross-check; Section B in the digest; options-gate exemption with NO CHAIN YET flag.
  \item \textbf{News catalysts:} Finnhub company news per universe ticker, NewsAPI secondary, keyword scoring, URL dedup; Section D in the digest. Strict quota budgeting (skip enrichment sources when near limits).
  \item \textbf{Macro event strip:} Static FOMC/CPI/PPI/jobs-report calendar (published yearly, stored as a local YAML, refreshed manually each quarter); rendered as a banner in the digest header.
  \item \textbf{Corporate actions:} Ex-dividend dates and stock splits via yfinance for universe tickers.
  \item \textbf{Lockup expirations:} Computed from IPO listing dates (+90/+180 days), persisted in a small local file as IPOs are observed.
  \item \textbf{OPEX markers:} Deterministic third-Friday monthly options expiration dates rendered in the header.
  \item \textbf{Scheduling:} cron (or Task Scheduler) at 6:00 AM ET with failure notification (non-zero exit writes an error HTML so the user sees something went wrong, not just a missing file). Resolve Open Question 5.
\end{enumerate}
\textbf{Exit criterion:} All four event categories live; scheduled run executes unattended for 5 trading days; quota usage stays within free tiers daily.

\subsection{Phase 3: Usability \& Delivery \hfill {\small (target: as needed)}}

\begin{enumerate}
  \item \textbf{Watchlist customization UI:} Simple local web page (or CLI) to edit \texttt{tickers.txt} and thresholds without opening files by hand.
  \item \textbf{Email delivery:} Send the daily digest via free SMTP (e.g., Gmail app password); HTML email variant of the template.
  \item \textbf{Sector filtering:} Sector tags per ticker (yfinance/Finnhub profile) with client-side filter buttons in the HTML digest.
  \item \textbf{Historical archive view:} Index page linking past digests; optional simple hit-rate review (did flagged events occur as scheduled), strictly factual, still no prediction.
\end{enumerate}
\textbf{Exit criterion:} User-defined; Phase 3 items ship independently as needed.

\subsection{Cross-Phase Engineering Rules}
\begin{itemize}
  \item Every step lands with unit tests (pytest); fetchers are tested against recorded fixture responses, not live APIs.
  \item No step may introduce a new critical data type with only one source (Constraint, Section 15).
  \item All output text in the digest is reviewed against the events-only principle: factual descriptions, no predictive or directional language.
  \item Secrets never enter the repo; config references environment variables only.
\end{itemize}

%=====================================================================
\section{MVP Scope}

\begin{callout}{MVP Goal}
A working daily Python script that generates a clean HTML digest with earnings + 8-K events, merged from redundant free sources and filtered by options volume. Run manually each morning. (Maps to Phase 1 of the Implementation Plan, Section~\ref{sec:impl}.)
\end{callout}

\subsection{In Scope, MVP}
\begin{itemize}
  \item Earnings calendar: Finnhub + yfinance union merge with confirmation status \hfill [expanded in v1.1]
  \item SEC 8-K filing fetch (EDGAR free API, Finnhub fallback)
  \item Options volume filter (yfinance primary, Polygon fallback)
  \item Market cap classification (yfinance primary, Finnhub fallback)
  \item Elevated options activity detection (20-day average comparison) + raw put/call ratio display
  \item Watchlist config file (\texttt{tickers.txt}) always included in the universe
  \item Graceful degradation, local caching, \texttt{seen\_events.json} dedup \hfill [NEW in v1.1]
  \item HTML digest output, one file per day, with diagnostics footer
  \item Manual execution (run the script each morning)
\end{itemize}

\subsection{Out of Scope, MVP (Phase 2+)}
\begin{itemize}
  \item IPO calendar (Finnhub + EDGAR cross-check), Phase 2
  \item News catalyst scoring (Finnhub / NewsAPI / Alpha Vantage), Phase 2
  \item Macro event header strip: FOMC, CPI/PPI, jobs report dates, Phase 2 \hfill [NEW idea in v1.1]
  \item Ex-dividend dates and stock splits via yfinance, Phase 2 \hfill [NEW idea in v1.1]
  \item IPO lockup expiration tracking (listing date + 90/180 days), Phase 2 \hfill [NEW idea in v1.1]
  \item Monthly options expiration (OPEX) date markers, Phase 2 \hfill [NEW idea in v1.1]
  \item Automated scheduling (cron job), Phase 2
  \item Watchlist customisation UI, Phase 3
  \item Email delivery of the digest, Phase 3
  \item Sector-based filtering UI, Phase 3
\end{itemize}

%=====================================================================
\section{Constraints \& Assumptions}

\subsection{Constraints}
\begin{itemize}
  \item \textbf{Zero cost:} No paid APIs, no SaaS subscriptions, no data licenses under any circumstance.
  \item \textbf{US equities only:} NYSE and NASDAQ listed securities. No OTC, no ADRs, no ETF-specific events in v1.
  \item \textbf{No single-source dependency:} Every critical data type must have a defined fallback (Section~\ref{sec:chains}). A new data type cannot ship with only one source. \hfill [NEW in v1.1]
  \item \textbf{Delayed options data:} Free-tier options data from Polygon.io is 15-minute delayed. Sufficient for next-day planning but not intraday.
  \item \textbf{yfinance reliability:} yfinance scrapes Yahoo Finance unofficially. It may break, skip tickers, or have thinner coverage outside widely held names. No SLA. This is the primary motivation for the redundancy strategy.
  \item \textbf{SEC EDGAR policy:} Must send a valid User-Agent header. Stay at or below 10 req/sec.
\end{itemize}

\subsection{Assumptions}
\begin{itemize}
  \item The user runs the script once per morning (Phase 1), moving to scheduled runs in Phase 2.
  \item A Python 3.10+ environment is available locally.
  \item Free API keys for Finnhub and Polygon.io have been obtained by the user (both free to sign up).
  \item The output HTML file is opened locally in a browser, no web server required.
  \item Stocks with no options chain are not of interest and can be safely excluded (IPO exception noted in Section 6.2).
\end{itemize}

%=====================================================================
\section{Success Metrics}

\begin{tabularx}{\textwidth}{|X|L{6.6cm}|}
\hline
\textbf{Metric} & \textbf{Target} \\
\hline
Earnings events captured (next 7 days) & $>$ 95\% of actual S\&P 500 earnings, measured against the union of sources \hfill [raised in v1.1] \\
\hline
Single-source blind spots & 0 critical data types served by only one source; coverage diff logged daily \hfill [NEW in v1.1] \\
\hline
False positive rate (noise events) & $<$ 20\% of digest entries feel irrelevant \\
\hline
Estimated-date mislabeling & 0 ESTIMATED earnings dates displayed as CONFIRMED \hfill [NEW in v1.1] \\
\hline
Options gate accuracy & 0 illiquid / micro-cap stocks pass through \\
\hline
Resilience & Digest still generates (partial, with banner) when any single source is fully down \hfill [NEW in v1.1] \\
\hline
Script runtime & $<$ 3 minutes for full daily run (with caching) \\
\hline
API quota usage & Stays within free-tier limits every day \\
\hline
Digest readability & All key fields visible without horizontal scroll in browser \\
\hline
\end{tabularx}

%=====================================================================
\section{Open Questions \hfill {\small [UPDATED in v1.1]}}

\subsection{Resolved in v1.1}

\begin{tabularx}{\textwidth}{|L{0.6cm}|X|L{5.6cm}|}
\hline
\textbf{\#} & \textbf{Question} & \textbf{Decision} \\
\hline
1 & Watchlist: hardcoded list or config file? & Config file (\texttt{tickers.txt} / \texttt{config.yaml}). Never hardcode. \\
\hline
2 & Show all stocks passing the gate, or watchlist + top unknowns? & Watchlist always shown, plus top N non-watchlist by rank score; N configurable (default 25). \\
\hline
4 & Separate legal events from M\&A in the digest? & Merged in MVP; split later only if volume justifies it. \\
\hline
\end{tabularx}

\subsection{Still Open}

\begin{tabularx}{\textwidth}{|L{0.6cm}|X|L{4.2cm}|}
\hline
\textbf{\#} & \textbf{Question} & \textbf{Decision Needed By} \\
\hline
3 & Is 500-contract options volume the right threshold, or per-cap-tier (lower for small caps, e.g., 200)? Leaning per-tier so the SMALL tag is not dead weight. & During Phase 1, Step 11 (validation week) \\
\hline
5 & Phase 2 scheduling: cron, or Claude Cowork scheduled tasks? & Phase 2 planning \\
\hline
6 & When earnings sources disagree on a date by more than 2 days, show both dates or only the earliest? \hfill [NEW in v1.1] & During Phase 1, Step 11 \\
\hline
7 & Should PREVIOUSLY REPORTED events be shown collapsed or fully suppressed by default? \hfill [NEW in v1.1] & During Phase 1, Step 11 \\
\hline
\end{tabularx}

%=====================================================================
\section{Revision History}

\begin{tabularx}{\textwidth}{|L{1.4cm}|L{1.7cm}|L{1.7cm}|X|}
\hline
\textbf{Version} & \textbf{Date} & \textbf{Author} & \textbf{Summary} \\
\hline
1.0 & Jun 2026 & Shreyas & Initial draft, requirements discussion with Claude Cowork \\
\hline
1.1 & Jun 2026 & Shreyas & Events-only principle formalized (directional labels removed, raw put/call ratio retained). Added: ticker universe definition, multi-source redundancy strategy with primary/fallback chains and union merge, earnings confirmation status, graceful degradation, local caching, cross-day event dedup, diagnostics footer, phased implementation plan (Section~\ref{sec:impl}), Phase 2 event ideas (macro calendar, ex-div/splits, lockup expirations, OPEX). Corrected Polygon.io free-tier rate limit. Resolved open questions 1, 2, 4. Document converted to LaTeX as the master format. \\
\hline
\end{tabularx}

\vspace{1.5em}
\begin{center}
\textit{End of Document, Market Event Radar PRD v1.1}
\end{center}

\end{document}
